---
title: "Midterm (75 minutes)"
author: "Polina"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=F,message = F,error = T)

```

```{r}
3+"as"
```

## Intro (2 points)

Solve the tasks, knit to html and upload the knitted document to
eclassrom. Set RMarkdown global setting to show the code, hide messages
and warnings and enable knitting even if the errors occur. If you will
upload only rmd file (not knitted to html), a half points will be
deducted from your final score.

Good luck!

\## Task 1 (5 points)

Import the data from:
<https://raw.githubusercontent.com/urosgodnov/datasets/master/diabetes.csv>

Show:

\- dataframe dimension - use sapply to show the class of every column -
convert Age column to factor with breakes (0,30],
(30,40],(40,60],(60,90] - show the frequencies of Age column - what is
the frequency of Age in the interval (40,60]?

```{r}
library(tidyverse)
"https://raw.githubusercontent.com/urosgodnov/datasets/master/diabetes.csv" %>% read.csv->df1
df1%>%dim()
df1%>%sapply(class)
df1$Age %>% cut(c(0,30,60,90))->df1$Age
df1$Age %>% summary()
df1$Age %>% summary() %>% .["(40,60)"] ##????
```

\## Task 2 (5 points)

Install and load janeaustenr package. With austen_books() function store
the data into dftext variable.

Answer:

\- how many different books are in the dftext variable?

\- how many rows does the book Persuasion contain?

\- how many rows contain the letter a/A?

\- how many rows start AND end with the letter a/A?

```{r}
library(janeaustenr)
dftext <- austen_books() 
dftext$book %>% unique() %>% length()
dftext %>% .[.$book=="Persuasion",] %>% nrow()
(dftext$book=="Persuasion") %>% sum()
dftext %>% .[grep("a|A",.$text),] %>% nrow()
dftext %>% .[grepl("(a|A)$",.$text) & grepl("^(a|A)",.$text),]
dftext %>% .[grepl("^(a|A).*(a|A)$",.$text),] %>% nrow()

```

\## Task 3 (4 points)

Create 100 random dates in the years 2021.

\- which month has the highest frequency? What is the frequency?

\- Which weekday has the highest frequency? What is the frequency?

\- choose the first 5 dates and print the whole weekday name in german
language.

```{r}
library(lubridate)
dates<-make_date(2021,sample(1:12),sample(1:28, 100, replace=TRUE))
"1.1.2021" %>% dmy() %>% {.+days(0:364)} %>% sample(100)
dates %>% month(label = T, abbr=F) %>% summary %>% .[max(.)==.]
dates %>% wday(label = T, abbr=F) %>% summary %>% .[max(.)==.]
dates %>% .[1:5] %>% wday(label=T,abbr=F, locate="German")
```

\## Task 4 (4 points)

Import data from
<https://raw.githubusercontent.com/urosgodnov/datasets/master/RandomData.csv>
into variable dfCar.

Answer:

\- what is the average cars' age (in years) in 2021?

\- what is the average (in days) from the purchase on today's date? Mind
the purchase date has the format m/d/y. Transform the result to seconds
with a proper function.

\- both results should be presented in a table 2x2, which you can create
with the Rmarkdown visual editor.

```{r}
"https://raw.githubusercontent.com/urosgodnov/datasets/master/RandomData.csv" %>% read.csv()->df4
df4 %>% head()
df4$CarModelYear %>% {2021-.} %>% mean()->a
df4$PurchaseDate %>% mdy() %>% difftime(today(),.,units="secs") %>% mean()->b
```

| task | result |
|------|--------|
| 1    | `r a` years  |
| 2    | `r b`  sec |

: Result

\## Task 5 (5 points)

Given a vector of numbers x\<-10001:20000 find all numbers, where sqrt
of a number is a whole number. Result should be one vector with all the
correct numbers!! How many numbers are in the result vector?

```{r}
x<-10001:20000
x %>% {s<-sqrt(.);.[round(s)==s]}
x %>% {s<-sqrt(.);.[round(s)==s]} %>% length()
```

